---
title: "Predicting IMBD Scores"
author: "Shivani Dedhia, Akhila Pamukuntla, Nafis Chowdhury, Akshita Jain"
output: 
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE , dev = "png", dpi=300)
options(digits = 2)

library(readr)
library(dplyr)
library(ggplot2)
library(tidyverse)
library(stringr)
library(ggrepel)
library(reshape2)
library(corrgram)
library(tree)
library(rpart)
library(rpart.plot)
library(randomForest)
library(plotly)
```

# STA 9750 Final Project

## Introduction

###### Many factors make a movie succesful and profitable. Some of the factors include budget of the movie, actor's and director's popularity, etc. We will find out which factor has the most impact on the movies success and profitability.  

###### We have the IMDB 5000 movie dataset, which has 28 variables, 5043 movies across 100 years in 66 countries. There are many varibles in the dataset such as Director, Actors, Duration, Gross, Budget, Genres, Facebook Likes, etc. 

```{r combine and clean data, warning=FALSE, include=FALSE}
IMDB <- read.csv("IMDM_ratings.csv")

imdb <- IMDB[!duplicated(IMDB), ]
imdb <- imdb[!is.na(imdb$gross), ]
imdb <- imdb[!is.na(imdb$budget), ]

imdb <- subset(imdb, select = -c(aspect_ratio))
imdb <- subset(imdb, select = -c(color))
imdb <- subset(imdb, select = -c(movie_imdb_link))
imdb <- subset(imdb, select = -c(language))

# Cleaning the movie title by removing the special character "Â" at the end and some whitespaces.
imdb$movie_title <- gsub("Â", "", as.character(factor(imdb$movie_title)))
str_trim(imdb$movie_title, side = "right")

# Adding profit and ROI
imdb <- imdb %>% mutate(net_profit = gross - budget,return_on_investment = (net_profit/budget)*100)

# replacing all NA's with col average
imdb$facenumber_in_poster[is.na(imdb$facenumber_in_poster)] <- round(mean(imdb$facenumber_in_poster, na.rm = TRUE))

# replacing all 0's with NA's
imdb[,c(5,6,8,13,24,26)][imdb[,c(5,6,8,13,24,26)] == 0] <- NA

# replacing all NA's with col average
imdb$num_critic_for_reviews[is.na(imdb$num_critic_for_reviews)] <- round(mean(imdb$num_critic_for_reviews, na.rm = TRUE))
imdb$duration[is.na(imdb$duration)] <- round(mean(imdb$duration, na.rm = TRUE))
imdb$director_facebook_likes[is.na(imdb$director_facebook_likes)] <- round(mean(imdb$director_facebook_likes, na.rm = TRUE))
imdb$actor_3_facebook_likes[is.na(imdb$actor_3_facebook_likes)] <- round(mean(imdb$actor_3_facebook_likes, na.rm = TRUE))
imdb$actor_1_facebook_likes[is.na(imdb$actor_1_facebook_likes)] <- round(mean(imdb$actor_1_facebook_likes, na.rm = TRUE))
imdb$cast_total_facebook_likes[is.na(imdb$cast_total_facebook_likes)] <- round(mean(imdb$cast_total_facebook_likes, na.rm = TRUE))
imdb$actor_2_facebook_likes[is.na(imdb$actor_2_facebook_likes)] <- round(mean(imdb$actor_2_facebook_likes, na.rm = TRUE))
imdb$movie_facebook_likes[is.na(imdb$movie_facebook_likes)] <- round(mean(imdb$movie_facebook_likes, na.rm = TRUE))

# delete the blank cols in content rating as they cannot be replaced with anything reasonable
imdb <- imdb[!(imdb$content_rating %in% ""),]

# replacing all content_rating with mordern rating system
imdb$content_rating[imdb$content_rating == 'M']   <- 'PG' 
imdb$content_rating[imdb$content_rating == 'GP']  <- 'PG' 
imdb$content_rating[imdb$content_rating == 'X']   <- 'NC-17'
imdb$content_rating[imdb$content_rating == 'Approved']  <- 'R' 
imdb$content_rating[imdb$content_rating == 'Not Rated'] <- 'R' 
imdb$content_rating[imdb$content_rating == 'Passed']    <- 'R' 
imdb$content_rating[imdb$content_rating == 'Unrated']   <- 'R' 
imdb$content_rating <- factor(imdb$content_rating)

levels(imdb$country) <- c(levels(imdb$country), "Others")
imdb$country[(imdb$country != 'USA')&(imdb$country != 'UK')] <- 'Others' 
imdb$country <- factor(imdb$country)
```

## Data Exploration

###### After exploring the data variables such as languages, aspect ratio, imdb movie link and color were removed to reduce bias. As many of these variables had data heavily skewed on one side. 

## Imdb score distribution

```{r visualizing, warning=FALSE, include=FALSE}

ggplot(aes(x=imdb_score), data = imdb) +
  geom_histogram(binwidth = 0.2,aes(fill = ..count..)) +
  scale_x_continuous(breaks = 0:10) +
  ggtitle("Imdb Score Distribution") +
  labs(x = "Imdb Score Distribution", y = "Number of Movies")

```
###### Count of movies rated above 7.5 which are considered to be highly recommended

###### Majority of the movies are rated 7.6 with only a handful rated above 9. The highest rating recieved by one movie is 9.3. This shows there is not enough data to understand the variables that have the strongest impact on highly rated movies.

```{r visualizing, warning=FALSE, include=FALSE}

imdb %>% group_by(imdb_score) %>% filter(imdb_score > 7.5) %>% count(imdb_score) %>% arrange(desc(n))

```
###### Content Rating Distribution

```{r visualizing, warning=FALSE, include=FALSE}

imdb %>% group_by(content_rating) %>%
  filter(!is.na(content_rating)) %>%
  summarise(average_imdb_score = mean(imdb_score), num = n()) %>%
  ggplot(aes(x=average_imdb_score, y = num, label = content_rating)) +
  geom_point() +
  geom_label_repel()

```
